### 2017-11-13 Draft2/Appending.Growth.tex
Make algorithms not float
Rename var

Let $\FuncWritesByGrow(n)$ denote the total number of writes made by $\FuncGrow$, and let $\VarNumItemsTail$ be the size of the tail. Since Corollary \ref{coro:GrowthSeq} also holds true for growth arrays, $\FuncGrow$ is called $|\VarGrowSeq|$ times. A buffer is appended to the tail each time $\FuncGrow$ is called. Thus, the tail's size is
\begin{align*}
\VarNumItemsTail &= |\VarGrowSeq| = |\VarCapacitySeq| - 1 = \VarUseful
\end{align*}
Then the formula for $\FuncWritesByGrow(n)$ is simply $\FuncWritesTail(\VarUseful)$, since $\VarUseful$ pointers are appended to the tail, which is a dynamic array. Finally, adding the writes made by $\FuncAppend$, the formula for $\FuncWrites(n)$ is
\begin{align*}
\FuncWrites(n) = n + \FuncWritesTail(\VarUseful)
\end{align*}
Now, I approximate $\FuncWrites(n)$ using $\sim$. To do this, I will derive the big-O complexity of $\VarUseful$.

@@

In the previous section, I established that $\VarNumItemsTail = \VarUseful$. Since $\VarCapacitySeq_\VarUseful$ is the capacity needed to hold $n$ items, and $\FuncSpaceTail(\VarUseful)$ is the space the tail needs to store $\VarUseful$ pointers, I conclude that the formula for $\FuncSpace(n)$ is
\begin{align*}
\FuncSpace(n) = \VarCapacitySeq_\VarUseful + \FuncSpaceTail(\VarUseful)
\end{align*}
($\FuncSpaceTail$ denotes the space cost function for dynamic arrays.)

>>>>

Let $\FuncWritesByGrow(n)$ denote the total number of writes made by $\FuncGrow$, and let $\VarNumPointersTail$ be the size of the tail. Since Corollary \ref{coro:GrowthSeq} also holds true for growth arrays, $\FuncGrow$ is called $|\VarGrowSeq|$ times. A buffer is appended to the tail each time $\FuncGrow$ is called. Thus, the tail's size is
\begin{align*}
\VarNumPointersTail &= |\VarGrowSeq| = |\VarCapacitySeq| - 1 = \VarUseful
\end{align*}
Then the formula for $\FuncWritesByGrow(n)$ is simply $\FuncWritesTail(\VarUseful)$, since $\VarUseful$ pointers are appended to the tail, which is a dynamic array. Finally, adding the writes made by $\FuncAppend$, the formula for $\FuncWrites(n)$ is
\begin{align*}
\FuncWrites(n) = n + \FuncWritesTail(\VarUseful)
\end{align*}
Now, I approximate $\FuncWrites(n)$ using $\sim$. To do this, I will derive the big-O complexity of $\VarUseful$.

@@

In the previous section, I established that $\VarNumPointersTail = \VarUseful$. Since $\VarCapacitySeq_\VarUseful$ is the capacity needed to hold $n$ items, and $\FuncSpaceTail(\VarUseful)$ is the space the tail needs to store $\VarUseful$ pointers, I conclude that the formula for $\FuncSpace(n)$ is
\begin{align*}
\FuncSpace(n) = \VarCapacitySeq_\VarUseful + \FuncSpaceTail(\VarUseful)
\end{align*}
($\FuncSpaceTail$ denotes the space cost function for dynamic arrays.)

### 2017-11-13 Draft2/Indexing.Growth.tex
Make algorithms not float
[WIP] Feedback
Rename var

\begin{algorithm}
	\begin{algorithmic}
		\Function{$\FuncGetItem$}{$\VarList,\ \ParamIndex$}
			\State $\LclIndex \gets \ParamIndex$
			\For{$\LclBuffer \TextIn \VarList.\FieldTail$}
				\If{$\LclIndex < \LclBuffer.\FieldLength$}
					\State \Return $\LclBuffer[\LclIndex]$
				\EndIf
				\State $\LclIndex \gets \LclIndex - \LclBuffer.\FieldLength$
			\EndFor
			\State \Return $\VarList.\FieldHead[\LclIndex]$
		\EndFunction
	\end{algorithmic}
\end{algorithm}

I know that this algorithm runs in $O(\log n)$ time since it was shown earlier that the size of the tail, $\VarNumItemsTail$, equals $\VarUseful$, and (from Lemma \ref{lem:VarUsefulIsOLogN}) that $\VarUseful = O(\log n)$. Aside from iteration over the tail, all statements run in constant time, so the time complexity is precisely $O(\log n)$.

@@

Now, consider that every buffer except the head is stored inside the tail. Thus, such buffers can be uniquely identified by their index in the tail. I will refer to this quantity as the \textbf{buffer index} and denote it $\VarIndexBuffer$. I wish to assign the head a buffer index as well, so that each buffer has a unique ID. Since the head follows the tail's last buffer, which has $\VarIndexBuffer = \VarNumItemsTail - 1$, I will let the head's $\VarIndexBuffer$ be $\VarNumItemsTail$.

I define a helper function, $\FuncGetBuffer$, that gets the buffer associated with a given $\VarIndexBuffer$. $\VarIndexBuffer$ is assumed to be valid; that is, $0 \leq \VarIndexBuffer \leq \VarNumItemsTail$.

\begin{algorithm}
	\begin{algorithmic}
		\Function{$\FuncGetBuffer$}{$\VarList,\ \LclIndexBuffer$}
			\If{$\LclIndexBuffer < \VarList.\FieldTail.\FieldSize$}
				\State \Return $\VarList.\FieldTail[\LclIndexBuffer]$
			\Else
				\State \Return $\VarList.\FieldHead$
			\EndIf
		\EndFunction
	\end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}
	\begin{algorithmic}
		\Function{$\FuncGetItem$}{$\VarList,\ \ParamIndex$}
			\State $(\LclIndexBuffer,\ \LclIndexElement) \gets \FuncDecompose(\ParamIndex)$
			\State \Return $\VarList.\FuncGetBuffer(\LclIndexBuffer)[\LclIndexElement]$
		\EndFunction
	\end{algorithmic}
\end{algorithm}

@@

	The last element always resides in the head, so $\VarIndexBuffer = \VarNumItemsTail|_{n = i + 1}$. As shown earlier, $\VarNumItemsTail = \VarUseful$, so $\VarIndexBuffer = \VarUseful|_{n = i + 1}$.

	To determine $\VarIndexElement$, consider the identity:
	\begin{align*}
	n = \text{\# of items in tail} + \text{\# of items in head}
	\end{align*}
	First, I will determine the number of items in the tail. In the case where $\VarIndexBuffer = 0$, then because $\VarNumItemsTail = \VarIndexBuffer$, the tail contains $0$ buffers and thus $0$ items (which equals $\VarGrowSeq_{-1}$). If $\VarIndexBuffer > 0$, the number of items in the tail is the last size at which $\FuncGrow$ was called, or the last term of $\VarGrowSeq$. This quantity is $\VarGrowSeq_{\VarUseful - 1}$, or $\VarGrowSeq_{\VarIndexBuffer - 1}$.

@@

\begin{algorithm}
	\begin{algorithmic}
		\Function{$\FuncDecompose$}{$\ParamIndex$}
			\State $\LclIndexBuffer \gets \max(\left\lceil \log_{\VarGrowthFactor} (\ParamIndex + 1) - \log_{\VarGrowthFactor} \VarInitCapacity \right\rceil, 0)$
			\If{$\VarIndexBuffer > 0$}
				\State $\LclIndexElement \gets \ParamIndex - \VarGrowthFactor^{\LclIndexBuffer - 1} \times \VarInitCapacity$
			\Else
				\State $\VarIndexElement \gets \ParamIndex$
			\EndIf
			\State \Return $(\LclIndexBuffer,\ \LclIndexElement)$
		\EndFunction
	\end{algorithmic}
\end{algorithm}

% Todo: CITE!
Clearly, $\FuncDecompose$ runs in constant time. Despite that, it appears to be quite expensive: normally, logarithms and exponentiation require use of floating-point instructions, which are slower than integer-based instructions. However, in the special case where $\VarGrowthFactor = 2$ and $\VarInitCapacity = 2^\VarLogInitCapacity$ for some constant whole number $\VarLogInitCapacity$, I claim that $\VarIndexBuffer$ and $\VarIndexElement$ can be computed without use of floating-point instructions.

>>>>

\begin{algorithm}[H]
	\begin{algorithmic}
		\Function{$\FuncGetItem$}{$\VarList,\ \ParamIndex$}
			\State $\LclIndex \gets \ParamIndex$
			\For{$\LclBuffer \TextIn \VarList.\FieldTail$}
				\If{$\LclIndex < \LclBuffer.\FieldLength$}
					\State \Return $\LclBuffer[\LclIndex]$
				\EndIf
				\State $\LclIndex \gets \LclIndex - \LclBuffer.\FieldLength$
			\EndFor
			\State \Return $\VarList.\FieldHead[\LclIndex]$
		\EndFunction
	\end{algorithmic}
\end{algorithm}

I know that this algorithm runs in $O(\log n)$ time since it was shown earlier that the size of the tail, $\VarNumPointersTail$, equals $\VarUseful$, and (from Lemma \ref{lem:VarUsefulIsOLogN}) that $\VarUseful = O(\log n)$. Aside from iteration over the tail, all statements run in constant time, so the time complexity is precisely $O(\log n)$.

@@

Now, consider that every buffer except the head is stored inside the tail. Thus, such buffers can be uniquely identified by their index in the tail. I will refer to this quantity as the \textbf{buffer index} and denote it $\VarIndexBuffer$. I wish to assign the head a buffer index as well, so that each buffer has a unique ID. Since the head follows the tail's last buffer, which has $\VarIndexBuffer = \VarNumPointersTail - 1$, I will let the head's $\VarIndexBuffer$ be $\VarNumPointersTail$.

I define a helper function, $\FuncGetBuffer$, that gets the buffer associated with a given $\VarIndexBuffer$. $\VarIndexBuffer$ is assumed to be valid; that is, $0 \leq \VarIndexBuffer \leq \VarNumPointersTail$.

\begin{algorithm}[H]
	\begin{algorithmic}
		\Function{$\FuncGetBuffer$}{$\VarList,\ \LclIndexBuffer$}
			\If{$\LclIndexBuffer < \VarList.\FieldTail.\FieldSize$}
				\State \Return $\VarList.\FieldTail[\LclIndexBuffer]$
			\Else
				\State \Return $\VarList.\FieldHead$
			\EndIf
		\EndFunction
	\end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}[H]
	\begin{algorithmic}
		\Function{$\FuncGetItem$}{$\VarList,\ \ParamIndex$}
			\State $(\LclIndexBuffer,\ \LclIndexElement) \gets \FuncDecompose(\ParamIndex)$
			\State \Return $\VarList.\FuncGetBuffer(\LclIndexBuffer)[\LclIndexElement]$
		\EndFunction
	\end{algorithmic}
\end{algorithm}

@@

	The last element always resides in the head, so $\VarIndexBuffer = \VarNumPointersTail|_{n = i + 1}$. As shown earlier, $\VarNumPointersTail = \VarUseful$, so $\VarIndexBuffer = \VarUseful|_{n = i + 1}$.

	To determine $\VarIndexElement$, consider the identity:
	\begin{align*}
	n = \text{\# of items in tail} + \text{\# of items in head}
	\end{align*}
	First, I will determine the number of items in the tail. In the case where $\VarIndexBuffer = 0$, then because $\VarNumPointersTail = \VarIndexBuffer$, the tail contains $0$ buffers and thus $0$ items (which equals $\VarGrowSeq_{-1}$). If $\VarIndexBuffer > 0$, the number of items in the tail is the last size at which $\FuncGrow$ was called, or the last term of $\VarGrowSeq$. This quantity is $\VarGrowSeq_{\VarUseful - 1}$, or $\VarGrowSeq_{\VarIndexBuffer - 1}$.

@@

\begin{algorithm}[H]
	\begin{algorithmic}
		\Function{$\FuncDecompose$}{$\ParamIndex$}
			\State $\LclIndexBuffer \gets \max(\left\lceil \log_{\VarGrowthFactor} (\ParamIndex + 1) - \log_{\VarGrowthFactor} \VarInitCapacity \right\rceil, 0)$
			\If{$\VarIndexBuffer > 0$}
				\State $\LclIndexElement \gets \ParamIndex - \VarGrowthFactor^{\LclIndexBuffer - 1} \times \VarInitCapacity$
			\Else
				\State $\VarIndexElement \gets \ParamIndex$
			\EndIf
			\State \Return $(\LclIndexBuffer,\ \LclIndexElement)$
		\EndFunction
	\end{algorithmic}
\end{algorithm}

Clearly, $\FuncDecompose$ runs in constant time. Despite that, it appears to be quite expensive: normally, logarithms and exponentiation require use of floating-point instructions, which are slower than integer-based instructions. However, in the special case where $\VarGrowthFactor = 2$ and $\VarInitCapacity = 2^\VarLogInitCapacity$ for some constant whole number $\VarLogInitCapacity$, I claim that $\VarIndexBuffer$ and $\VarIndexElement$ can be computed without use of floating-point instructions.
